Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study

### 2025-05-06 docs/response/response.qmd

---
title: "Response to Journal's Reviewer Comments"
subtitle: "Conditionally Autoregressive Models for House Price Data: Insights From a Comparative Simulation Study"
---

**Reviewer 1**

**This manuscript compares different methods (Conditionally Autoregressive, Geographically Weighted Regression, and Simultaneous Autoregressive) for modeling property prices. The problem definition looks reasonable, but the manuscript presents only the application part; nothing is contributed toward the theory. Therefore, I will suggest transferring this manuscript to an application-related journal. To my understanding, this manuscript does not fall in the scope of this journal. Further, the manuscript is not standard format of the journal.**

We acknowledge that our work is primarily application-driven, focusing on the use of Conditionally Autoregressive (CAR), Geographically Weighted Regression (GWR), and Simultaneous Autoregressive (SAR) models in the context of property price modeling. However, we respectfully submit that the manuscript makes a meaningful methodological contribution, particularly in bridging the gap between spatial theory and applied practice—an area of growing relevance, especially in spatial analysis for developing regions.

While we do not introduce new theoretical constructs, our work extends the methodological discourse in the following ways:

 - **Shifting the Dominant Modeling Paradigm**: Previous studies on house price modeling have predominantly employed Geographically Weighted Regression (GWR), which requires precise geolocation data for model fitting. However, such data are often incomplete, unavailable, or unreliable in many regions. Our study explores the potential of the Conditionally Autoregressive (CAR) model—more commonly applied in health and ecological studies—as a viable alternative for house price modeling. This cross-domain adaptation is underexplored and opens new methodological pathways for spatial econometrics.

 - **Unified Comparative Framework**: We develop and apply a systematic framework to compare CAR, SAR, and GWR models using both simulated data and real-world property data. This comparative approach, particularly within the under-researched housing market context of a developing region, fills a critical gap in the literature.

 - **Practical Implementation Insights**: The manuscript offers practical guidance for applied researchers on implementing, diagnosing, and interpreting each spatial model. These details are especially valuable for practitioners operating in data-scarce environments or policy-driven contexts.

 - **Informing Model Selection**: By revealing the comparative strengths and limitations of each model in capturing spatial heterogeneity, our findings contribute to informed model selection in applied spatial econometrics—balancing methodological soundness with practical feasibility.

 - **Policy and Planning Relevance**: The results of our study have implications for spatial planning and housing market analysis, offering a framework for integrating spatial models into decision-making processes where traditional approaches may be constrained by data availability.

Given these points, we believe the manuscript fits within the journal’s broader mission to advance rigorous spatial analysis methods and encourage cross-disciplinary methodological applications.

**Reviewer 2**

1.  **The choice of $W$ (whether binary adjacency or row-standardized) is critical. Have the authors explored alternative spatial weight matrices, and how do they impact model performance?**

    Thank you for this insightful comment regarding the specification of the spatial weight matrix, $W$. We fully acknowledge that the choice of $W$—whether binary adjacency, row-standardized, or distance-based—can substantially influence the performance and interpretation of spatial models. In the present study, our primary objective was to evaluate and compare the performance of CAR, SAR, and GWR models using both simulated data and real-world data from Lombok. To maintain consistency across models and to isolate the structural differences in model behavior, we have, at this stage, limited our analysis to a binary adjacency matrix for $W$. We agree that exploring alternative specifications of $W$ (e.g., row-standardized or distance-decay matrices) is a valuable direction for further research. We plan to extend our analysis in future work to assess how varying spatial weight structures affect the robustness and predictive accuracy of these spatial models.

2.  **While the CAR model shows lower bias and RMSE in the artificial study, does this advantage persist in real-world scenarios, particularly where data sparsity or boundary effects exist?**

    Thank you for your thoughtful question. We acknowledge that in small-area data, bias can indeed become a concern due to limited information at the unit level. However, in the case of the multilevel CAR (mlvCAR) model, each area contains multiple observations, allowing us to model area-specific effects—functionally resembling a random effects structure. This hierarchical structure helps mitigate the impact of sparsity within individual areas.

    Regarding sample size, prior studies suggest that having approximately 15–30 observations per area is generally sufficient for stable estimation in multilevel frameworks. Given our data structure aligns with this recommendation, we believe the mlvCAR model offers a robust approach, especially in real-world scenarios where heterogeneity and partial pooling across areas are important. We thus recommend the use of mlvCAR to address both the potential bias and boundary effects effectively.

3.  **The estimated value of $\rho$ (spatial dependence strength) varies widely (e.g., 0.48 for the multilevel CAR model). Have the authors examined how sensitive the model results are to changes in $\rho$?**

    In our study, we recognize that the value of $\rho$ (the spatial dependence parameter) plays a crucial role in determining the extent to which spatial autocorrelation is captured by the model. While we did not conduct a dedicated sensitivity analysis by fixing $\rho$ at multiple values, our simulation framework implicitly accounts for variability in $\rho$ through repeated data generation and model fitting, where $\rho$ is treated as an estimable parameter. The observed variability—such as the estimate of 0.48 in the multilevel CAR model—reflects the differences in underlying spatial structure across datasets rather than instability or inconsistency in the model itself.

    In a Bayesian framework, the prior distribution assigned to $\rho$ also has an important influence, particularly in data-sparse contexts or when spatial signals are weak. In our implementation, we used a uniform or weakly informative prior (depending on the model setting), allowing the data to drive the estimation of $\rho$ while still preventing extreme or implausible values. In practical applications, however, researchers may wish to incorporate more informative priors based on prior knowledge of spatial processes or domain-specific expectations, which can enhance stability and interpretability in certain contexts.

    That said, we fully agree that a more systematic sensitivity analysis—such as fixing $\rho$ at various plausible levels or experimenting with different prior distributions—could offer deeper insights into model robustness and the influence of spatial structure. We consider this an important direction for future research, particularly in settings with heterogeneous spatial patterns or limited sample sizes, where sensitivity to $\rho$ may significantly impact conclusions and policy relevance.

4.  **The study discusses the spatial effects $\phi$ as a key advantage of the CAR model, but how can these effects be meaningfully interpreted for house price determinants? Would spatial effect maps or additional diagnostic tools enhance this understanding?**

    We agree that while $\phi$ captures unobserved spatial heterogeneity, its interpretability in the context of house price determinants requires careful consideration. Rather than directly reflecting specific covariates, $\phi$ accounts for spatially structured residual variation—such as omitted neighborhood characteristics, local amenities, or spatial spillovers—that are not captured by observed variables.

    For a brief illustration, since the model uses log-transformed prices, $\phi$ has multiplicative interpretation on the original price scale. So if $\phi = 0.57$ for a given sub-district, then after accounting for covariates, properties in that area are estimated to be approximately $\exp(0.57) \approx 1.76$ or 76% more expensive than average. Similarly, a sub-district with $\phi = -0.2$, this suggests that properties in that area are approximately 18% less expensive than average, since $\exp(-0.2) \approx 0.82$. Note that the 95% credible interval of $\phi$ can be interpreted similarly by exponentiating the lower and upper bounds, yielding a multiplicative uncertainty range for the price deviation.

    To facilitate interpretation, we provide a map that visualizes the estimated spatial effects ($\phi$), thereby highlighting the spatial patterns identified by the model.

5.  **Among the various spatial regression models, Geographically Weighted Regression (GWR) is particularly prominent in property price research (page 2). Before reaching this conclusion, did you compare GWR to other spatial regression models?**

    In our manuscript, we highlight Geographically Weighted Regression (GWR) as a prominent method in the literature, particularly due to its frequent application in house price studies. This observation is based on a review of previous works rather than a performance-based comparison, for instance @wang2020geographically, @lu2011geographically, @stewart2018localized, and @cellmer2020spatial. While GWR is indeed widely used for its ability to model spatial heterogeneity at the local level, our study goes further by systematically comparing GWR with CAR and SAR models through simulation and real data applications. This comparative analysis aims to evaluate their relative strengths and limitations more objectively.

6.  **The scope of properties analysed includes residential houses and commercial real estate (page 16). How do the authors reconcile this with the abstract, which mentions only house price data?**

     We acknowledge that the description in the manuscript may not have been sufficiently clear. The original dataset obtained through web-scraping did include a broad range of property types—such as residential houses, villas, boarding houses, and vacant land. However, for the purposes of this study, we restricted our analysis to residential houses only. We will revise the manuscript accordingly to ensure that this scope is clearly stated in both the abstract and the data collection section.

7.  **After data cleaning, the final dataset included 598 entries representing multiple observations per sub-district (page 16). Why does the distribution of these data not appear on the map (Figure 6)?**

    For clarity, we note that the Lombok house price dataset used in this study does not contain precise geographic coordinates for each observation. As a result, it is not possible to accurately display the distribution of the data points on a map. While each observation is linked to a sub-district, the absence of point-level location data limits our ability to visualize them spatially within those areas. Therefore, for a more informative visualization, we are preparing a heatmap of house prices at the sub-district level, which will reflect the spatial variation in prices across Lombok using aggregated data.

8.  **The authors have applied point-level data using three different models (GLMM, GWR, and multilevel CAR) (page 17). How was this approach implemented, given that there is no explanation of how each observation was georeferenced?**

    Thank you for this important observation. We clarify that GLMM and multilevel CAR models do not require explicit geographic coordinates for each individual observation. Instead, they rely on an area-level identifier (ID) that links each observation to a corresponding spatial unit (e.g., sub-district). This is a similar approach to random effects or multilevel modelling.

    In contrast, GWR requires spatial coordinates for each observation. To meet this requirement, we generated random points around the centroid of each sub-district, using `st_sample()`, where the number of points corresponds to the number of observations in that area. Each point was placed within a 0.5 km radius from the centroid, ensuring a reasonable spatial spread while maintaining the area-specific context. This approach enabled us to simulate realistic spatial positions for property transactions in the absence of precise geolocation data. Note that, the same bandwith was used in the fitting of the GWR model. By enforcing the area-level grouping, this ensures a fair comparison to the previous method.

9.  **Why is Table 5 not referred to in the text?**

    In the revised version of the manuscript, we have made sure to explicitly refer to Table 5 in the main text to improve clarity and consistency. We appreciate your careful reading and helpful suggestion.

10. **The authors have argued that point-level models like GWR are limited when precise location data are unavailable (page 18). However, for the Lombok House Price Data, the authors have used point-level models. How do you reconcile this contradiction?**

    As we noted in our response to Comment 8, the Lombok house price data do not come with precise geographic coordinates. In order to apply the `gwr()` function, which requires point-level location data, we generated synthetic coordinates by simulating random points around the centroid of each sub-district. While this workaround allowed us to run the GWR model, we acknowledge that it is not an ideal or practical solution for real-world applications. This limitation reinforces our argument that point-level models such as GWR can be challenging to apply meaningfully when precise location data are not available.

11. **In the Lombok house price analysis, no sensitivity analysis was conducted to test how different definitions of adjacency (W matrix) influence the results. Would the findings remain consistent if a different spatial weights structure was used?**

    As mentioned in our response to Comment 1, in this study we used a binary adjacency matrix to define spatial relationships, primarily to maintain consistency across the CAR and SAR models and to focus on comparing their general performance. We acknowledge that the choice of spatial weights matrix ($W$) can significantly influence model outcomes. While we have not conducted a sensitivity analysis in this current version, we agree that testing alternative specifications—such as row-standardized or distance-based matrices—would be valuable. We consider this an important direction for future work to assess the robustness of our findings under different definitions of spatial connectivity.

12. **The study primarily evaluates CAR models within the Lombok context, but how well does it generalize to larger, high-density urban areas (e.g., Jakarta)? Would additional testing in a more complex metropolitan setting strengthen the validity of the conclusions?**

    Thank you for this important point. Indeed, numerous studies have already explored property price modeling in large, high-density urban areas such as Jakarta, where spatial data availability and heterogeneity are typically greater. In contrast, there is relatively limited research focused on low-density or semi-rural contexts such as Lombok. Our study aims to address this gap by evaluating the applicability and performance of spatial models, particularly the CAR model, in a setting that has different data characteristics and spatial structures compared to major metropolitan areas.

    We also acknowledge that in richer datasets—common in urban contexts where more granular, high-frequency, and multidimensional data are available—machine learning approaches may offer competitive or even superior predictive performance. However, our focus here is to assess interpretable spatial models in contexts where data limitations remain a challenge.

# References

### 2025-05-06 docs/_tex/manuscript.tex
update manuscript and response

Recent studies have leveraged GWR to investigate the spatial
heterogeneity of housing market determinants. For example, Wang and Chen
(2020) applied GWR to examine the impact of local built-environment
factors on home prices across different phases of the housing market
cycle. Their findings highlighted significant spatial variations,
underscoring the critical role of local context in influencing home
prices. Similarly, Lu, Charlton, and Fotheringham (2011) employed GWR
with a non-Euclidean distance metric to analyze London house price data.
By incorporating this alternative metric, they better captured the
spatially varying relationships between house prices and their
determinants, emphasizing the importance of considering different
distance metrics in spatial analyses. In their 2018 study, Stewart
Fotheringham and Park (2018) explored spatial and temporal variations in
the determinants of apartment prices in Seoul, South Korea, over a
decade. Utilizing a hedonic price model with a spatio-temporal lag,
calibrated through GWR, they revealed consistent spatial variations and
strong spatial lag effects on house prices over time. Additionally,
Cellmer, Cichulska, and Bełej (2020) applied GWR to assess the spatial
variations in determinants of the housing market in Poland. The study
identified spatial patterns of autocorrelation in average housing prices
and market activity, illustrating the value of GWR in examining
localized effects of socio-demographic and environmental factors on the
housing market. Collectively, these studies demonstrate the efficacy of
GWR and its extensions in capturing the nuanced spatial and temporal
variations in housing market data, providing valuable insights for urban
planning and policy-making.

@@

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

@@

\[
\begin{array}{cccccc}
    & A & B & C & D & E \\
A & 0 & 1 & 1 & 1 & 1\\ 
B & 1 & 0 & 1 & 0 & 0\\ 
C & 1 & 1 & 0 & 1 & 0\\
D & 1 & 0 & 1 & 0 & 1\\
E & 1 & 0 & 0 & 1 & 0\\
\end{array}
\]

@@

For instance, in SAR and CAR models, the diagonal matrix
\(\boldsymbol{D}\) aids in incorporating spatial lag and error
components by appropriately weighting the influence of neighbouring
regions (Wall (2004), Ver Hoef et al. (2018)). In GWR, this matrix
assists in locally calibrating the model by reflecting the density and
connectivity of regions (Brunsdon, Fotheringham, and Charlton (1996),
Stewart Fotheringham and Park (2018)). Here, \(\boldsymbol{D}\) is
represented as:

\[
\begin{array}{cccccc}
   & A & B & C & D & E \\
A & 4 & 0 & 0 & 0 & 0\\ 
B & 0 & 2 & 0 & 0 & 0\\ 
C & 0 & 0 & 3 & 0 & 0\\
D & 0 & 0 & 0 & 3 & 0\\
E & 0 & 0 & 0 & 0 & 2\\
\end{array}
\]

This matrix format indicates that region \(A\) has 4 neighbour, regions
\(C\) and \(D\) each have 3 neighbours, and region \(B\) and \(E\) has 2
neighbour.

@@

\includegraphics[width=0.7\textwidth,height=\textheight]{figures/nb_artfcountry.png}

@@

Here, \(\boldsymbol{\varepsilon} \sim N(0, \nu^2)\) while
\(\boldsymbol{\phi}\) is a specific component in CAR model that has a
role as spatial effect. It is also common to mention
\(\boldsymbol{\phi}\) as a CAR priors. It is a type of Gaussian Markov
random field (Rue and Held (2005)), capture spatial autocorrelation by
ensuring that values at nearby locations are more similar than those
further apart (Lee (2013)). This can be expressed in a general term

\begin{equation}\phantomsection\label{eq-phidist}{
\boldsymbol{\phi}\sim N(0,\tau^2\boldsymbol{Q}^{-1}) 
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-phicar}{
\phi_k|\boldsymbol\phi_{-K}, \boldsymbol{W},\rho, \tau^2  \sim  N\left(\frac{\rho\sum_{i=1}^Kw_{ki}\phi_i}{A},\frac{\tau^2}{A}\right)
}\end{equation}

@@

The Bayesian approach to CAR models entails defining prior distributions
for all model parameters, such as the regression coefficients \(\beta\),
variance parameters \(\nu^2\) and \(\tau^2\), and spatial random effects
\(\phi_k\). Subsequently, Bayesian inference techniques like Markov
Chain Monte Carlo (MCMC) are employed to derive posterior distributions
of these parameters. This method facilitates the integration of prior
knowledge and offers a versatile framework for quantifying uncertainty.

@@

\toprule
Elements & $GWR$ & $SAR$ & $CAR$\\
\midrule
Number of Observations ($N$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Covariate Matrix ($X$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Spatial Weights Matrix ($W$) & $\times$ & $\checkmark$ & $\checkmark$\\
Spatial Coordinates (coords) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Covariate Coefficients ($\beta$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\addlinespace
Error Term variance ($\sigma^2$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Spatial Autocorrelation ($\rho$) & $\times$ & $\checkmark$ & $\checkmark$\\
Spatial Structure component ($\phi$) & $\times$ & $\times$ & $\checkmark$\\
Spatial structure variance ($\tau^2$) & $\times$ & $\times$ & $\checkmark$\\
Additional scale ($\nu^2$) & $\times$ & $\times$ & $\checkmark$\\
\addlinespace
Estimation method & MLE & MLE/GMM & MCMC\\
\bottomrule

@@

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

By employing artificially generated covariates and an artificially
constructed study region, we developed three distinct property price
datasets, each based on a different spatial model: CAR, a combination of
CAR and SAR, and GWR. The true parameter values for this purpose are
\(\boldsymbol{\beta} = \begin{bmatrix} 1 & 0.9 & 0.7 & 0.5 & 0.3 \end{bmatrix}\),
\(\nu^2 = 0.2\), \(\tau^2 = 0.6\) and \(\rho = 0.6\). Whereas beta
values for GWR dataset is generated using a beta function, which is a
function of coordinate points. The matrix \(\mathbf{W}\) was derived
from the neighbourhood structure of an artificial study region using the
functions \texttt{poly2nb()} and \texttt{nb2mat()}. Data generation for
\(SAR + CAR\) employs a slightly different strategy. In addition to
being generated per area simultaneously using the joint distribution of
\(\mathbf{Y}\) as in Equation~\ref{eq-jointysar}, information about the
spatial effect \(\boldsymbol{\phi}\) is also added at the end of the
process. Therefore, we refer to it as \(SAR + CAR\) data.

@@

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

Lombok includes five districts (kota/kabupaten), 53 sub-districts
(kecamatan), and 608 villages (desa/kelurahan), which represent three
successive levels of administrative division. In this study, the
sub-district level is used as the spatial unit of analysis, as
illustrated in Figure~\ref{fig-lombokmap}. The key variables of interest
include property price, land size, built-up area, number of bedrooms,
and number of bathrooms.

@@

Further, we conducted several preprocessing steps. First, we filtered
the dataset by removing irrelevant variables, including floors,
furnishing status, and category. It was then subsequently filtered to
retain only properties with plausible characteristics: land area between
90 and 800 square meters, built-up size between 70 and 600 square
meters, a maximum of 6 bedrooms, and a maximum 5 bathrooms. These
thresholds were applied to ensure the data reflect realistic and
context-appropriate housing characteristics in Lombok, based on common
residential patterns and local housing norms. This step was essential to
enhance the validity of the analysis by excluding outliers or
potentially erroneous entries. Following this procedure, the dataset was
refined to comprise a total of 576 observations. A summary of the key
variables is presented in Table~\ref{tbl-datasumm}.

@@

\begin{longtable}[t]{lrrrrrrrr}

\caption{\label{tbl-datasumm}Summary statistics of variables in the
property price dataset}

@@

\toprule
Variable & Mean & Median & $Q_1$ & $Q_3$ & Min & Max & SD & NA\\
\midrule
Prices (million IDR) & 1983.00 & 1400 & 800.0 & 2500.0 & 170 & 22500 & 2056.81 & 26\\
Land-size (sqm) & 299.60 & 255 & 147.2 & 407.2 & 90 & 800 & 179.39 & 24\\
Built-up area (sqm) & 184.80 & 150 & 100.0 & 210.0 & 70 & 600 & 109.20 & 24\\
No. of bedroom & 3.24 & 3 & 2.0 & 4.0 & 1 & 6 & 1.19 & 24\\
No. of bathroom & 2.55 & 2 & 2.0 & 3.0 & 1 & 5 & 1.05 & 24\\
\bottomrule

@@

To address missing values, we employed the \texttt{mice} package (Van
Buuren and Groothuis-Oudshoorn (2011)) to perform multiple imputation
with parameters set to \(maxit = 10\), \(m = 10\), and a fixed random
seed to ensure reproducibility. The imputed values were then
reintegrated into the main dataset, replacing the original missing
entries. This process resulted in a final dataset that was complete and
ready for subsequent analysis using the selected spatial models.

@@

@@

The cleaned and completed Lombok dataset was then model. The tables
below present the results of the data fitting, offering a comparative
view of each model's effectiveness and accuracy in capturing the
dataset's characteristics.

\begin{equation}\phantomsection\label{eq-lopreg}{
prices \sim land_{std} + building_{std} + beds_{std} + baths_{std}
}\end{equation}

Standardization of the predictors ensures that the resulting
coefficients are on a comparable scale, facilitating meaningful
interpretation and comparison across different modeling approaches. The
results presented in Table~\ref{tbl-modelcomp} provide a comprehensive
evaluation of each model's performance in capturing both the structural
and spatial characteristics of the property market in Lombok.

We conducted model fitting on the area-level data, with detailed results
provided in the Appendix (Table~\ref{tbl-modelcomp1}). However, this
method does not provide optimal parameter estimates because data
aggregation can reduce variability and obscure spatial details essential
for capturing local effects. To address this, we used point-level data
to apply three different models: GLMM, GWR, and multilevel CAR.

@@

\textsubscript{Source:
\href{https://indiraputeri-phd.github.io/CAR_simcomp/manuscript.qmd.html}{Article
Notebook}}

>>>>

Recent studies have leveraged GWR to investigate the spatial
heterogeneity of housing market determinants. For example, C.-H. Wang
and Chen (2020) applied GWR to examine the impact of local
built-environment factors on home prices across different phases of the
housing market cycle. Their findings highlighted significant spatial
variations, underscoring the critical role of local context in
influencing home prices. Similarly, Lu, Charlton, and Fotheringham
(2011) employed GWR with a non-Euclidean distance metric to analyze
London house price data. By incorporating this alternative metric, they
better captured the spatially varying relationships between house prices
and their determinants, emphasizing the importance of considering
different distance metrics in spatial analyses. In their 2018 study,
Stewart Fotheringham and Park (2018) explored spatial and temporal
variations in the determinants of apartment prices in Seoul, South
Korea, over a decade. Utilizing a hedonic price model with a
spatio-temporal lag, calibrated through GWR, they revealed consistent
spatial variations and strong spatial lag effects on house prices over
time. Additionally, Cellmer, Cichulska, and Bełej (2020) applied GWR to
assess the spatial variations in determinants of the housing market in
Poland. The study identified spatial patterns of autocorrelation in
average housing prices and market activity, illustrating the value of
GWR in examining localized effects of socio-demographic and
environmental factors on the housing market. Collectively, these studies
demonstrate the efficacy of GWR and its extensions in capturing the
nuanced spatial and temporal variations in housing market data,
providing valuable insights for urban planning and policy-making.

@@

@@

\[
\mathbf{W} =
\begin{array}{cccccc}
   & A & B & C & D & E \\
A & 0 & 1 & 1 & 1 & 1\\ 
B & 1 & 0 & 1 & 0 & 0\\ 
C & 1 & 1 & 0 & 1 & 0\\
D & 1 & 0 & 1 & 0 & 1\\
E & 1 & 0 & 0 & 1 & 0\\
\end{array}
\hspace{2.5cm}
\mathbf{D} =
\begin{array}{cccccc}
   & A & B & C & D & E \\
A & 4 & 0 & 0 & 0 & 0\\ 
B & 0 & 2 & 0 & 0 & 0\\ 
C & 0 & 0 & 3 & 0 & 0\\
D & 0 & 0 & 0 & 3 & 0\\
E & 0 & 0 & 0 & 0 & 2\\
\end{array}
\]

@@

For instance, in SAR and CAR models, the diagonal matrix
\(\boldsymbol{D}\) aids in incorporating spatial lag and error
components by appropriately weighting the influence of neighbouring
regions (Wall (2004), Ver Hoef et al. (2018)). In GWR, this matrix
assists in locally calibrating the model by reflecting the density and
connectivity of regions (Brunsdon, Fotheringham, and Charlton (1996),
Stewart Fotheringham and Park (2018)). In this case, \(\mathbf{D}\)
indicates that region \(A\) has 4 neighbours, regions \(C\) and \(D\)
each have 3 neighbours, and region \(B\) and \(E\) has 2 neighbours.

@@

\includegraphics[width=0.6\textwidth,height=\textheight]{figures/nb_artfcountry.png}

@@

Here, \(\boldsymbol{\varepsilon} \sim N(0, \sigma_{\varepsilon}^2)\)
while \(\boldsymbol{\phi}\) is a specific component in CAR model that
has a role as spatial effect. It is also common to mention
\(\boldsymbol{\phi}\) as a CAR priors. It is a type of Gaussian Markov
random field (Rue and Held (2005)), capture spatial autocorrelation by
ensuring that values at nearby locations are more similar than those
further apart (Lee (2013)). This can be expressed in a general term

\begin{equation}\phantomsection\label{eq-phidist}{
\boldsymbol{\phi}\sim N(0,\sigma_{\phi}^2\boldsymbol{Q}^{-1}) 
}\end{equation}

@@

\begin{equation}\phantomsection\label{eq-phicar}{
\phi_k|\boldsymbol{\phi}_{-k}, \mathbf{W},\rho, \sigma_{\phi}^2  \sim  N\left(\frac{\rho\sum_{i=1}^Kw_{ki}\phi_i}{A},\frac{\sigma_{\phi}^2}{A}\right)
}\end{equation}

@@

The Bayesian approach to CAR models entails defining prior distributions
for all model parameters, such as the regression coefficients \(\beta\),
variance parameters \(\sigma_{\varepsilon}^2\) and \(\sigma_{\phi}^2\),
and spatial random effects \(\phi_k\). Subsequently, Bayesian inference
techniques like Markov Chain Monte Carlo (MCMC) are employed to derive
posterior distributions of these parameters. This method facilitates the
integration of prior knowledge and offers a versatile framework for
quantifying uncertainty.

@@

\toprule
Elements & $GWR$ & $SAR$ & $CAR$\\
\midrule
Number of Observations ($N$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Covariate Matrix ($X$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Spatial Weights Matrix ($W$) & $\times$ & $\checkmark$ & $\checkmark$\\
Spatial Coordinates (coords) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Covariate Coefficients ($\beta$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
\addlinespace
Spatial Autocorrelation ($\rho$) & $\times$ & $\checkmark$ & $\checkmark$\\
Spatial Structure component ($\phi$) & $\times$ & $\times$ & $\checkmark$\\
Spatial structure variance ($\sigma_{\phi}^2$) & $\times$ & $\times$ & $\checkmark$\\
Error Term variance ($\sigma_{\varepsilon}^2$) & $\checkmark$ & $\checkmark$ & $\checkmark$\\
Estimation method & MLE & MLE/GMM & MCMC\\
\bottomrule

@@

By employing artificially generated covariates and an artificially
constructed study region, we developed three distinct property price
datasets, each based on a different spatial model: CAR, a combination of
CAR and SAR, and GWR. The true parameter values for this purpose are
\(\boldsymbol{\beta} = \begin{bmatrix} 1 & 0.9 & 0.7 & 0.5 & 0.3 \end{bmatrix}\),
\(\sigma_{\phi}^2 = 0.2\), \(\sigma_{\phi}^2 = 0.6\) and \(\rho = 0.6\).
Whereas beta values for GWR dataset is generated using a beta function,
which is a function of coordinate points. The matrix \(\mathbf{W}\) was
derived from the neighbourhood structure of an artificial study region
using the functions \texttt{poly2nb()} and \texttt{nb2mat()}. Data
generation for \(SAR + CAR\) employs a slightly different strategy. In
addition to being generated per area simultaneously using the joint
distribution of \(\mathbf{Y}\) as in Equation~\ref{eq-jointysar},
information about the spatial effect \(\boldsymbol{\phi}\) is also added
at the end of the process. Therefore, we refer to it as \(SAR + CAR\)
data.

@@

Lombok comprises five districts (kota/kabupaten), 53 sub-districts
(kecamatan), and 608 villages (desa/kelurahan), representing three
successive levels of administrative division. Mataram District, despite
its small area compared to other districts in Lombok, is the most
urbanized and economically active area on the island, characterized by
dense residential clusters with relatively homogeneous housing
specifications. In adjacent districts such as West Lombok and Central
Lombok, housing developments have expanded, particularly to accommodate
commuters working in Mataram. The coastal areas of West Lombok,
including Batu Layar and Sekotong, exhibit villa-style housing commonly
associated with tourism, resulting in generally higher property prices.
A similar pattern is found in North Lombok, where mountainous terrain
limits development, but tourism-driven demand has led to villa-style
housing near popular destinations. East Lombok, by contrast, remains
more isolated due to its distance from Mataram. Housing in this district
tends to serve primarily the local population, with limited signs of
external investment or tourism-driven development. Additionally, the
southern part of Central Lombok (Pujut) has seen emerging residential
and real estate activity following the construction of the international
MotoGP circuit.

In this study, the sub-district level is used as the spatial unit of
analysis, as illustrated in Figure~\ref{fig-lombokmap}. The key
variables of interest include property price, land size, built-up area,
number of bedrooms, and number of bathrooms.

@@

Further, we conducted several preprocessing steps. First, we filtered
the dataset by removing irrelevant variables, including floors,
furnishing status, and category. It was then subsequently filtered to
retain only properties with plausible characteristics: land area between
90 and 800 square meters, built-up size between 70 and 600 square
meters, a maximum of 6 bedrooms, and a maximum 5 bathrooms. These
thresholds were applied to ensure the data reflect realistic and
context-appropriate housing characteristics in Lombok, based on common
residential patterns and local housing norms. This step was essential to
enhance the validity of the analysis by excluding outliers or
potentially erroneous entries. Following this procedure, the dataset was
refined to comprise a total of 576 observations. A summary of the key
variables is presented in Table~\ref{tbl-datasumm}. Notably, the
distribution of property prices is skewed and does not exhibit a
bell-shaped curve, which motivates the inclusion of a log-transformed
price variable in the analysis.

@@

\begin{longtable}[t]{lrrrrrrr}

\caption{\label{tbl-datasumm}Summary statistics of house prices dataset
(\(N = 576\), \texttt{NA} = 4\%), including variables such as prices and
its logscale value, land area(sqm), building area(sqm), number of beds,
and number of baths}

@@

\toprule
Variable & Mean & SD & Median & $Q_1$ & $Q_3$ & Min & Max\\
\midrule
Prices (million IDR) & 1979.20 & 2054.51 & 1400.00 & 800.00 & 2500.00 & 170.00 & 22500.00\\
logPrices & 7.26 & 0.80 & 7.24 & 6.68 & 7.82 & 5.14 & 10.02\\
Land-size (sqm) & 299.08 & 179.33 & 254.50 & 147.00 & 405.75 & 90.00 & 800.00\\
Built-up area (sqm) & 184.46 & 109.14 & 150.00 & 100.00 & 210.00 & 70.00 & 600.00\\
No. of bedroom & 3.23 & 1.19 & 3.00 & 2.00 & 4.00 & 1.00 & 6.00\\
\addlinespace
No. of bathroom & 2.54 & 1.05 & 2.00 & 2.00 & 3.00 & 1.00 & 5.00\\
\bottomrule

@@

To address missing values, we used the \texttt{mice} package (Van Buuren
and Groothuis-Oudshoorn (2011)), which performs multiple imputation by
chained equations (MICE). This approach iteratively fills in missing
data by modeling each variable with missing values as a function of
other variables in the dataset. By generating multiple imputed datasets,
it accounts for the uncertainty inherent in missing data. In our case,
we specified 10 iterations and generated 10 imputed datasets. A fixed
random seed was used to ensure the results were reproducible. The
imputed datasets were then pooled to obtain final estimates for
subsequent analyses. The imputed values were then reintegrated into the
main dataset, replacing the original missing entries. This process
resulted in a final dataset that was complete and ready for subsequent
analysis using the selected spatial models.

@@

Moreover, as GWR requires spatial coordinates for each observation, we
addressed this by generating random points around the centroid of each
sub-district using the \texttt{st\_sample()} function. The number of
points generated matched the number of observations in each area, and
each was positioned within a 0.5 km radius of the centroid to maintain a
realistic spatial distribution while preserving the local context. This
strategy allowed us to simulate plausible spatial locations for property
transactions in the absence of precise geolocation data. Importantly, we
applied the same bandwidth in the GWR model fitting to ensure
consistency. By preserving the area-level grouping structure, this
approach supports a fair comparison with the previous method.

@@

The cleaned and completed Lombok dataset was then fitted to the model.
Equation Equation~\ref{eq-lopreg} presents the proposed model
specification, which serves as the fixed formula applied consistently
across all analyses

\begin{equation}\phantomsection\label{eq-lopreg}{
\text{logprices} \sim \text{land}_{\text{std}} + \text{building}_{\text{std}} + \text{beds}_{\text{std}} + \text{baths}_{\text{std}}
}\end{equation}

Standardization of the predictors ensures that the resulting
coefficients are on a comparable scale, facilitating meaningful
interpretation and comparison across different modeling approaches. We
conducted model fitting on the area-level data, with detailed results
provided in the Appendix (Table~\ref{tbl-modelcomp1}). However, this
method does not provide optimal parameter estimates because data
aggregation can reduce variability and obscure spatial details essential
for capturing local effects. To address this, we used point-level data
to apply three different models: GLMM, GWR, and multilevel CAR.

The results presented in Table~\ref{tbl-modelcomp} provide a
comprehensive evaluation of each model's performance in capturing both
the structural and spatial characteristics of the property market in
Lombok.

@@

Here, Table~\ref{tbl-modelcomp} compares parameter estimates and fit
criteria for three models: GLMM, GWR, and multilevel CAR. Significant
coefficients are marked with an asterisk (*). Key determinants like
land-size and built-up area have positive, significant effects across
models, while the number of bedrooms shows only minor, varying effects,
indicating model-specific differences in capturing this variable's
influence. Only the multilevel CAR model includes the spatial parameter
\(\rho\), with a significant estimate of 0.48, showing the model's
ability to account for spatial dependence. Variance components
\(\sigma_{\varepsilon}^2\) (residual) and \(\sigma_{\phi}^2\) are also
detailed, with both terms are significant. Fit criteria (AIC, WAIC, and
Log-likelihood) suggest mlvCAR and GWR may better capture local spatial
variation, with mlvCAR balancing spatial structure and precision.
